---
title: "Landmarks defining transforms"
author: "Gregory Jefferis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Landmarks defining transforms}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Intro
One of the core features of elmr is to enable data in the FAFB EM space to be
transformed into one of a number of standard light level template brains. This
transformation depends on manually selected landmarks. We can inspect these
landmarks and to try to measure their consistency.

### Setup
```{r, message=FALSE}
library(elmr)
library(knitr)
opts_chunk$set(fig.width=5, fig.height=5)
# set up for 3d plots based on rgl package
rgl::setupKnitr()
# frontal view
view3d(userMatrix=rgl::rotationMatrix(angle = pi, 1,0,0), zoom=0.7)
```

## Landmarks
The standard landmarks used by elmr were specified by Davi Bock on the FAFB and
JFRC2013 brains using the [elm](https://github.com/saalfeldlab/elm) tool
developed by Stefan Saalfeld and John Bogovic. We can plot the landmarks in the
context of the JFRC2013 brain as follows.

```{r, rgl=TRUE}
plot3d(JFRC2013)
# note that the landmarks are in raw voxel coordinates and must be scaled
xyz=scale(elm.landmarks[,c("X","Y","Z")], scale = 1/voxdims(JFRC2013), center = FALSE)
xyz=as.data.frame(xyz)
rownames(xyz)=sub("Pt-","",elm.landmarks$Label)
spheres3d(xyz, col = ifelse(elm.landmarks$Use, "green", 'red'), radius = 4)
```

## Landmarks based transform
The standard `elm.landmarks` set is used within to the elmr package to register
a transformation that can map FAFB->JFRC2013 space in conjunction with the 
`xform_brain` functions provided by the `nat.templatebrains` packages.

Let's first look at the FAFB landmarks

```{r}
# note that these positions are in nm units and do not need scaling
xyz.fafb=elm.landmarks[,c("X1","Y1","Z1")]
rownames(xyz.fafb)=sub("Pt-","",elm.landmarks$Label)
kable(head(xyz.fafb))
```

Now we can try to look at the consistency of the transform between each landmark
pair as follows:

1. Drop out each landmark pair in turn
2. Calculate position of that landmark in target space using remaining landmark pairs
3. Measure difference between calculated position and manually defined position

```{r}
# record transformed position with leave one out
xyzt.loo=xyz
for(i in 1:nrow(xyz)) {
  thisreg <- reglist(tpsreg(xyz.fafb[-i, ], xyz[-i, ]))
  xyzt.loo[i,]=xform(xyz.fafb[i, , drop=FALSE], thisreg)
}
deltas=sqrt(rowSums((xyzt.loo-xyz)^2))
hist(deltas)
```

Let's looks at the deltas (distance between leave one out vs manually defined landmark position)
for each point.

```{r}
kable(cbind(elm.landmarks, delta=deltas))
```

And then visualise them in 3D, colouring each point by the magnitude of the delta.

```{r, rgl=TRUE, fig.width=8, fig.height=8}
clear3d()
view3d(userMatrix=rgl::rotationMatrix(angle = pi, 1,0,0), zoom=0.7)
jet.colors <-
  colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan",
                     "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

plot3d(JFRC2013)
spheres3d(xyz, col = jet.colors(10)[cut(deltas, 10)], radius = 4)
# nb these are numbered from 0 like the elm landmarks
texts3d(xyz, texts = rownames(xyz), adj = c(1,1))
```

Based on this, I would definitely suggest checking the following landmarks

75/48
43
46
68
23

Compare the leave one out landmark position with manually defined position

```{r, rgl=TRUE}
clear3d()
view3d(userMatrix=rgl::rotationMatrix(angle = pi, 1,0,0), zoom=0.7)
for(i in 1:nrow(xyz)){
  segments3d(rbind(xyz[i, , drop=FALSE], xyzt.loo[i, , drop=FALSE]), col=jet.colors(10)[cut(deltas, 10)][i], lwd=3)
}
points3d(xyz)
texts3d(xyz, texts = rownames(xyz), adj = c(1,1))
plot3d(JFRC2013)
```

Now of course this is not just a measure of consistency in each landmark point, 
but it will also depend strongly on the extent to which the neighbouring points
can define the required registration; in areas requiring substantial non-rigid
deformation this may be a problem; this will be further emphasised when the
neighbouring points are distant. We can check the latter point like so:

```{r}
library(nabor)
nearest3=rowMeans(nabor::knn(xyz, k=4)$nn.dists[,-1])
plot(deltas~nearest3)
deltabynearest3=lm(deltas~nearest3)
abline(deltabynearest3)
summary(deltabynearest3)
```

So there is a correlation between neighbour distance and our calculated delta;
with an R^2 of about 0.4 it's reasonably strong. This may be explained by
Davi placing more landmarks in areas where he is more certain.

Perhaps the ideal thing would be to find a set of landmarks on one side of the
brain and then find exactly the same landmarks on the other side of the brain.
For the JFRC2013 brain we can exactly mirror points from one side to the other 
using an image-based registration distributed with the `nat.flybrains` package.

We can find the mirror image position of all the points that we used like so:

```{r, rgl=TRUE}
#  nb this is conditioned on being able to find CMTK
if(isTRUE(nzchar(cmtk.bindir()))){
  xyzm=mirror_brain(xyz, brain = JFRC2013)
  
  clear3d()
  spheres3d(xyz, col='red', rad=2)
  spheres3d(xyzm, col='green', rad=2)
  plot3d(JFRC2013)
  view3d(userMatrix=rgl::rotationMatrix(angle = pi, 1,0,0), zoom=0.7)

}
```

One could then write out a table of landmarks including these mirrored positions
like so:

```{r, eval=FALSE}
xyzm.pixels=scale(xyzm, scale = voxdims(JFRC2013), center = FALSE)
colnames(xyzm.pixels)=c("XM","YM","ZM")
write.table(cbind(elm.landmarks, xyzm.pixels), 'elm.landmarks-with-mirror.tsv',
            col.names = T, sep='\t', row.names = F)
```


If one were then to take the mirror image landmarks in JFRC2013 space and 
manually pick the corresponding locations in EM space, one should be able to 
compare two different paths to the same point (one mirroring in FAFB, the other
mirroring in JFRC2013) and see whether there is a difference in the calculated
position. Differences would presumably reflect the uncertainty in picking 
positions in the EM volume (and a small contribution from the mirroring 
uncertainty in JFRC2013, which nevertheless expect to be <1 µm based on analysis in
[Manton et al 2014](https://doi.org/10.1101/006353)).
