---
title: "Making Surface Meshes"
author: "Gregory Jefferis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making surface meshes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Intro
You can use 3D data pulled from CATMAID to create meshes. You can also do the 
same with light level data and then transform into FAFB space.

### Setup
First load main packages
```{r, message=FALSE}
library(elmr)
cl=try(catmaid_login())
catmaid_available=inherits(cl, "catmaid_connection")
library(knitr)
opts_chunk$set(eval=catmaid_available)
```

In order to run some of this example we need to ensure that we:

* can login to the FAFB catmaid server

We will most our examples run conditionally based on this.

```{r, message=FALSE}
library(dplyr)
library(alphashape3d)
library(rgl)
# set up for 3d plots based on rgl package
rgl::setupKnitr()
# frontal view
view3d(userMatrix=rgl::rotationMatrix(angle = pi, 1,0,0), zoom=0.8)
```


## Selecting synapses 

We're going to use synapses between PNs and KCs to define a mesh

```{r, message=FALSE}
pntokc = catmaid_get_connectors_between(pre_skids = 'annotation:^PN$',
                                        post_skids = 'annotation:^KC$')
# select 3d position of presynaptic connectors
pntokc %>% 
  distinct(connector_id, .keep_all=TRUE) %>%
  select(connector_x:connector_z) %>%
  data.matrix -> cxyz
```

Now we can make an alphashape based on those connector positions. We need to set
the alpha value in the ashape3d function to something sensible based on the 
spatial scale of our data - he we use alpha=6000 nm.

```{r}
calyx.a=ashape3d(cxyz, alpha = 6000)
calyx=as.mesh3d(calyx.a)
```

We can compare the resultant mesh with the location of the connectors

```{r, rgl=TRUE}
wire3d(calyx)
points3d(cxyz, col='red')
points3d(pntokc[,c('post_node_x','post_node_y','post_node_z')], col='blue')
```

We can also compare with the mesh supplied with elmr

```{r, rgl=TRUE}
wire3d(calyx)
plot3d(FAFB13NP.surf, "MB_CA_R", alpha=.3)
view3d(userMatrix=rgl::rotationMatrix(angle = pi, 1,0,0), zoom=0.8)
```
And in dorsal view as well
```{r, rgl=TRUE}
wire3d(calyx)
plot3d(FAFB13NP.surf, "MB_CA_R", alpha=.3)
view3d(userMatrix=rgl::rotationMatrix(angle = -pi/2, 1,0,0), zoom=0.8)
```
