---
title: "Selecting random synapses"
author: "Gregory Jefferis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Selecting random synapses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Intro
You can select synapses within a spatial region and then carrying out a 
random sub-sampling. This might be useful as part of a strategy for defining
connectivity without reconstructing all connections.

### Setup
First load main packages
```{r}
library(elmr)
library(catmaid)
library(nat)
```

In order to run some of the examples we need to ensure that we:

* can login to the FAFB CATMAID server

We will most our examples run conditionally based on this.

```{r, message=FALSE}
cl=try(catmaid_login())
catmaid_available=inherits(cl, "catmaid_connection")
run_vignette=catmaid_available

# set up for 3d plots based on rgl package
rgl::setupKnitr()
# frontal view
view3d(userMatrix=rgl::rotationMatrix(angle = pi, 1,0,0), zoom=0.6)
```


## Selecting synapses 

We're going to look at synapses downstream of DA1 PNs

```{r, message=FALSE, eval=run_vignette}
cda1=catmaid_get_connectors_between('name:PN Glomerulus DA1')
```

If we wanted to get the upstream synapses, we could do that as follows:

```{r, message=FALSE, eval=FALSE}
cda1.in=catmaid_get_connectors_between(post_skids = 'name:PN Glomerulus DA1')
```

We then want to select synapses within a neuropil region, the lateral horn. The
`elmr` package now has a neuropil surface object `FAFBNP13.surf` that we can use
for this purpose. To look for points inside the right LH, we can extract just
that region from the surface object containing all neuropils as follows:

```{r}
# convert subset of regions to rgl::mesh3d object
lhr=as.mesh3d(FAFBNP.surf, 'LH_R')
```

Now we can select just those synapses within the LH volume, by piping the
XYZ positions of the presynaptic connector node to the `pointsinside` function.

```{r, eval=run_vignette}
library(dplyr)
# make a new logical column indicating if connector is inside LH
cda1$inlh=select(cda1, connector_x:connector_z) %>% 
  pointsinside(surf = lhr)
cda1 %>%
  filter(inlh) ->
  cda1.lh
```

Now let's take a random 10% sample of the post-synaptic connector nodes for just
one PN:

```{r, eval=run_vignette}
cda1.lh %>%
  filter(pre_skid==61221) %>%
  sample_frac(size=0.1) ->
random_syn
```

We can then construct CATMAID URLs for all those synapses:

```{r, eval=run_vignette}
synapse_urls=sapply(1:nrow(random_syn), 
       function(i) open_fafb(random_syn[i, c("post_node_x", "post_node_y", "post_node_z")],
            active_skeleton_id = random_syn$post_skid[i],
            active_node_id = random_syn$post_node_id[i], open = F)
)
```

That was a little bit clunky but we basically feed the `open_fafb` function the
data one row at a time centering the URL on the postsynaptic node and selecting
that node in the CATMAID viewer.

So here are the results

```{r, eval=run_vignette}
head(synapse_urls)
```

## Using the results

Let's make a plot to get an idea of where those synapses are:
First fetch the full structure of the seed PN

```{r, eval=catmaid_available}
seedpn=read.neuron.catmaid(61221)
```

Now plot the selected and unselected synapses 

```{r, eval=run_vignette}
# find bounding box containing selected synapses
synapse_bounds=apply(random_syn[,c("post_node_x", "post_node_y", "post_node_z")],2,range)
plot(seedpn, WithNodes=F, boundingbox=synapse_bounds)
# plot the synapses we didn't select
cda1.lh %>%
  filter(!post_node_id %in% random_syn$post_node_id & pre_skid %in% random_syn$pre_skid) %>%
  select(post_node_x:post_node_y) %>%
  points(col='grey')
# and then the ones we did
points(random_syn$post_node_x, random_syn$post_node_y, col='red', pch=19)
```

Then you can open those urls directly in CATMAID:

```{r, eval=FALSE}
for (u in synapse_urls) {
  readline("Press a key to move to open the next synapse (or Esc to cancel).")
  browseURL(u)
}
```

or write them to a text file

```{r, eval=FALSE}
writeLines(synapse_urls, con = 'PN61221_rand_synapse_urls.txt')
```

or upload them to a google doc

```{r, eval=FALSE}
library(googlesheets)
random_syn$url=synapse_urls
random_syn$tracer=""
random_syn$date_started=""
tf=tempfile(fileext = '.tsv')
write.table(random_syn, file=tf, sep="\t")
gs_upload(tf, sheet_title = 'PN61221_rand_synapse_urls')
unlink(tf)
```

