---
title: "Finding Neurons with Multiple Somata"
author: "Gregory Jefferis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
---
## Intro
Occasionally CATMAID users add the soma tag multiple times in error. We can
find such neurons and generate CATMAID URLs to inspect them.

### Setup
First load main packages
```{r, message=FALSE}
library(elmr)
cl=try(catmaid_login())
catmaid_available=inherits(cl, "catmaid_connection")
library(knitr)
# only run if catmaid available
# and cache so only run once per day
opts_chunk$set(eval=inherits(cl, "catmaid_connection"),cache.extra=Sys.Date())
library(dplyr)
rgl::setupKnitr()
```

## Finding neurons with multiple somata
First let's fetch information about all the labels (aka tags) applied to nodes
in the current project.
```{r, cache=FALSE}
label_stats=catmaid_get_label_stats()
```

Now let's restrict to cases where there are multiple soma tags per skeleton
```{r, munge soma tags}
# select soma labels
soma_labels = label_stats %>%
  filter(labelName == 'soma') %>%
  group_by(skeletonID)

# select skeleton ids for neurons with multiple cell bodies
multiple_soma = soma_labels %>%
  count(skeletonID) %>%
  filter(n > 1) %>%
  arrange(desc(n))
  multiple_soma_info = soma_labels %>%
  filter(skeletonID %in% multiple_soma$skeletonID)

multiple_soma_info = soma_labels %>% 
  filter(skeletonID%in% multiple_soma$skeletonID)
```

## XYZ position of the nodes we picked

```{r, cache=FALSE}
# wrapper for tree node details
node_details <- function(tnid) {
  res=catmaid_get_treenodes_detail(tnid)
  ul=catmaid_get_user_list()
  res$login=ul$login[match(res$user_id, ul$id)]
  # we expect these to be capitalised elsewhere
  cn=colnames(res)
  cn[cn%in%c("x","y","z")]=toupper(cn[cn%in%c("x","y","z")])
  colnames(res) <- cn
  res[c("X","Y","Z","radius","login")]
}

# note that we need to transpose the results of vnode_xyz to get X,Y,Z columns
multiple_soma_info <- cbind(as.data.frame(multiple_soma_info),
                            node_details(multiple_soma_info$treenodeID))
```

Now let's calculate the distance from the neuropil surface

```{r, neuropil distance}
multiple_soma_info$d=pointsinside(xyzmatrix(multiple_soma_info),
                                  FAFB.surf, rval = 'distance')
multiple_soma_info %>% 
  arrange(skeletonID, d) %>% 
  group_by(skeletonID) %>% 
  mutate(rank=row_number()) -> multiple_soma_info
kable(multiple_soma_info)
```

This lets us see that in some cases there are two soma tags outside the neuropil
(negative d) and close together - these are probably duplicates - whereas in 
other cases it is likely that points were added in error.
We can also plot the points colouring them by their rank order (most external 
first). 

```{r, rgl=TRUE, fig.width=8, fig.height=6}
# make a colour palette with as many entries as the maximum number of soma 
# tags in a neuron
pal=rainbow(max(multiple_soma_info$rank))
multiple_soma_info %>% 
  with(expr = spheres3d(X,Y,Z, col=pal[rank], rad=2000))
plot3d(FAFB)
par3d(zoom=.6)
```

Now we can use this information to construct an url for each node.

```{r, make urls}
multiple_soma_info %>%
  rowwise() %>%
  mutate(url = open_fafb(
    cbind(X, Y, Z),
    active_skeleton_id = skeletonID,
    active_node_id = treenodeID,
    open = FALSE
  )) -> multiple_soma_info

```

It might be useful to know who 'owns' each neuron. I think the
simplest way to assign this is by the user who has traced most nodes for each
skeleton (since there may be different users responsible for each soma).

```{r}
get_top_user <- function(x, ...) {
  ul=catmaid_get_user_list(...)
  # save time by checking unique skids only
  ddx=unique(x)
  gtu_one <- function(x, ...) {
    t <- try({
      res=catmaid_get_contributor_stats(x, ...)
      w=which.max(res$node_contributors$n)
      ul$login[match(res$node_contributors$id[w], ul$id)]
    })
    if(inherits(t, 'try-error')) NA_character_ else t
  }
  topus <- sapply(ddx, gtu_one)
  topus[match(x, ddx)]
}
```

## Google sheet
Let's make a google sheet with all those urls that we can then review manually:

```{r, googlesheet, eval=FALSE}
library(googlesheets)
# helper function to upload via temp file
# since writing cells via API is very slow
gs_upload_tf <- function(x, ...) {
  tf=tempfile(fileext = '.tsv')
  on.exit(unlink(tf))
  write.table(x, file=tf, sep="\t", row.names = FALSE)
  gs_upload(tf, ...)
}

multiple_soma_info %>% 
  arrange(skeletonID, d) %>% 
  group_by(skeletonID) %>%
  mutate(user=get_top_user(skeletonID)) %>%
  gs_upload_tf(sheet_title = 'multi_soma_neurons')
```

As an alternative we can divide that up with one worksheet per user. 

```{r, googlesheet-by-user, eval=FALSE}
library(googlesheets)
gs <- googlesheets::gs_new("multi_soma_neurons_by_user")
gs_add_sheet <- function(x, gs, ...) {
  gs_ws_new(row_extent = nrow(x)+1, col_extent = ncol(x), ss = gs, ..., input=x, col_names=T)
}
multiple_soma_info %>% 
  arrange(skeletonID, d) %>% 
  mutate(user=factor(get_top_user(skeletonID))) -> msi2

for(u in levels(msi2$user)) {
  gs_add_sheet(subset(msi2, user==u), gs, ws_title=u)
  cat(".")
}
```
